Implementation of Adaptive Waveform Inversion via a Penalty Function

### 2023-12-30 repro/papers/awigrad/paper.tex
near final

{\bf Note:}
$F$ is the mapping implemented (in finite difference approximation) in {\tt asg.fsbop}, with the additional constraint that the buoyancy field is fixed so that the computational version of ${\cal M}$ consists of (gridded) bulk modulus fields.

@@

The convolution operator implemented in the module {\tt segyvc} actually implements a time-discrete version of $K$ (or $L$). Note that the transpose $K[u]^T$ is a restricted version of cross-correlation, implemented as the transpose of {\tt segyvc.ConvolutionOperator}.

@@

{\bf Note:} IWAVE implements the mappings $R$ and $W$ in the form of
the commands {\tt iwave/trace/main/rms.x} and {\tt
  iwave/trace/main/txgain.x}. The natural python implementation of $R$
would output the vector of trace norms as a python array (or a NumPy
ndarray). However that would require extracting the number of traces
from the input, and checking that the inputs are compatible. This is
already implicitly done in the IWAVE manipulations of SU trace
data. So {\tt iwave/trace/main/rms.x} stores the vector of trace norms
as the 0th data sample in a set of traces, one for each trace of the
input data: all header words are the same except for {\tt ns}, which
is set to 1.

The python functions {\tt awi.setrms} and {\tt awi.applytxgain} call
the two IWAVE commands, manipulating the data as SU data files, rather
than as VCL vectors. That is, these are simply functions, not {\tt
  vcl.Function} objects. Since VCL vectors defined on a {\tt segyvc.Space}
have access to their data files (keyword {\tt data}), these are
convenient to use to express the action of $R$ and $W$ in AWI penalty operator,
gradient, and so on.

\section{Objectives}
Define for $\alpha \ge 0$
\[
 J_{\alpha,\sigma}[u,m,d] = \frac{1}{2}(\|S[m]u - d\|_D^2 + \alpha^2\|T_\sigma[m]u\|_U^2 + \sigma^2 \|u\|^2_U).
\]
Note that the notation is consistent, that is, reduces to $J_{0,\sigma}$ as defined above for $\alpha=0$.

@@

{\bf Note:} The implementation {\tt vcalg.conjgrad} of the Conjugate Gradient algorithm for the normal equations optionally returns the residual vector (parameter {\tt e}). The obvious way to use this is to create a 3-component product space for the output of the block-column operator $(S[m],\alpha T[m], \sigma I)^T$, and set the right-hand side vector (parameter {\tt b})
to $(d,0,0)^T$. Then on return one-half the norm-squared of {\tt e} is precisely the (approximate) value of $\tilde{J}_{\alpha,\sigma}[m]$.

@@

Then equation \ref{eqn:penderiv} can be re-written as
\[
  \langle D(T_{\sigma}[m]u)\delta m, T_{\sigma}[m]u\rangle = \langle DF[m]\delta m, 
\]
\begin{equation}
  \label{eqn:penderiv}
   K[v_{0,\sigma}[m,u]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u]\rangle
\end{equation}
Combine equations \ref{eqn:resderiv} and \ref{eqn:penderiv} with
equation \ref{eqn:basederiv} to obtain
\[
  \nabla \tilde{J}_{\alpha,\sigma}[m] =
  DF[m]^T\large\{ K[u_{\alpha,\sigma}[m]]^T (S[m]u_{\alpha,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:deriv}
  + \alpha^2 (
  K[v_{0,\sigma}[m, u_{\alpha,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+ K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{\alpha,\sigma}[m]]) \large\}.
\end{equation}

@@

The derivative of the expression on the right-hand side of
\ref{eqn:wg} is the sum of two terms, the first of which is given by
\ref{eqn:penderiv} with $u = u_{0,\sigma}[m]$. The second is
\[
  \langle T_{\sigma}[m]u_{0,\sigma}[m], T_{\sigma}[m]D_m
  u_{0,\sigma}[m]\delta m \rangle
\]
which from identity \ref{eqn:blockderiv} is
\[
  = -\langle  T_{\sigma}[m]^TT_{\sigma}[m]u_{0,\sigma}[m], (S[m]^TS[m] + \sigma^2I)^{-1}
\]
\[
  \times ((DS[m]^{T}\delta m) (S[m]u_{0,\sigma}[m]-d) +
  S[m]^T(DS[m]\delta) m u_{0,\sigma}[m])
\]
Since
$T_{\sigma}[m]^TT_{\sigma}[m] = W(Ru_{0,\sigma}[m],-2,2)$, this is
\[
  = -\langle (DS[m]\delta m) (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m],
  S[m]u_{0,\sigma}[m] - d \rangle
\]
\begin{equation}
  \label{eqn:term2}
  - \langle S[m] (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m], (DS[m]\delta
  m)u_{0,\sigma}[m] \rangle
\end{equation}
Introduce the abbreviation
\begin{equation}
  \label{eqn:defv1}
  v_{1,\sigma}[m] = (S[m]^TS[m] + \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m].
\end{equation}
Then the RHS of equation \ref{eqn:term2bis} may be re-written as
\begin{equation}
  \label{eqn:term2bis}
  = -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle.
\end{equation}
Recall that we have just computed the second of two terms, the sum of
which is the derivative of $\tilde{J}_{\sigma}[m]$. The first is
\ref{eqn:penderiv} with $u = u_{0,\sigma}[m]$. Combine the two
terms to obtain
\[
  D\tilde{J}_{\sigma}[m] \delta m = \langle DF[m]\delta m,
\]
\[
  K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{0,\sigma}[m]]\rangle
\]
\[
  -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle
\]
\[
  = \langle DF[m]\delta
  m,(K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]-K[v_{1,\sigma}[m]])^T(S[m]u_{0,\sigma}[m]-d)
\]
\[
  + K[u_{0,\sigma}[m]]^T S[m] (v_{0,\sigma}[m,u_{0,\sigma}[m]] -
  v_{1,\sigma}[m]) \rangle.
\]
To simplify this expression further, and reduce the computational
complexity of the derivative expression, note that the
operator-valued function $K$ is linear in its argument, according to
its definition \ref{eqn:convdef}. So the above is
\begin{equation}
  \label{eqn:wgderiv}
  = \langle DF[m]\delta m, (K[v_{\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d) 
  + K[u_{0,\sigma}[m]]^T S[m] v_{\sigma}[m] )\rangle.
\end{equation}
in which
\[
  v_{\sigma}[m] = v_{0,\sigma}[m,u_{0,\sigma}[m]] -  v_{1,\sigma}[m].
\]
The definitions \ref{eqn:defv0} and \ref{eqn:defv1} imply that
$v_{\sigma}[m]$ is the solution of the linear system
\begin{equation}
  \label{eqn:defvsigma}
  (S[m]^TS[m]+ \sigma^2I) v_{\sigma}[m] =
  (W(R(W(Ru_{0,\sigma}[m],-2,1)u_{0,\sigma}),2,0) - W(Ru_{0,\sigma}[m],-2,2))
  u_{0,\sigma}[m].
\end{equation}
That is, computation of $v_{\sigma}[m]$, hence of $D
\tilde{J}_{\sigma}[m]\delta m$, requires the solution of only one additional
linear system, beyond that necessary to compute the value of $\tilde{J}_{\sigma}[m]$.

@@

It remains to establish that the same result (expression
\ref{eqn:wgderiv}) for the derivative can be derived from the
limit \ref{eqn:wglim}. From from equation \ref{eqn:basederiv},
\[
  \frac{1}{\alpha^2}(D \tilde{J}_{\alpha,\sigma}[m] \delta m - D
  \tilde{J}_{0,\sigma}[m]\delta m) =
\]
\[
  = \frac{1}{\alpha^2} \left(\langle D_m(S[m]u)\delta m_{u =  u_{\alpha,\sigma}[m]}, (S[m]u_{\alpha,\sigma}[m]-d) \rangle\right.
\]
\[
  - \left.\langle D_m(S[m]u)\delta m_{u =  u_{0,\sigma}[m]}, (S[m]u_{0,\sigma}[m]-d) \rangle\right)
\]
\[
  +  \langle D_m(T_{\sigma}[m]u)\delta
  m|_{u=u_{\alpha,\sigma}[m]},T_{\sigma}[m]u_{\alpha,\sigma}[m]\rangle_U
\]
Since the normal operator \ref{eqn:normal} is uniformly positive
definite hence boundedly invertible uniformly in $\alpha \ge 0$,
$u_{\alpha,\sigma}[m] \rightarrow u_{0,\sigma}[m]$ in $U$. All of the
operators appearing in the third term are $U-$continuous, so  the
third term tends to
\[
  \langle D_m(T_{\sigma}[m]u)\delta
  m|_{u=u_{0,\sigma}[m]},T_{\sigma}[m]u_{0,\sigma}[m]\rangle_U
\]
which is identixal to the left-hand side of equation
\ref{eqn:penderiv} and is the first term in the direct computation of
the $\tilde{J}_{\sigma}$ derivative, analyzed above.

>>>>

@@

@@

\section{Objectives}
Define for $\alpha \ge 0$
\begin{equation}
  \label{eqn:j}
 J_{\alpha,\sigma}[u,m,d] = \frac{1}{2}(\|S[m]u - d\|_D^2 + \alpha^2\|T_\sigma[m]u\|_U^2 + \sigma^2 \|u\|^2_U).
\end{equation}
Note that the notation is consistent, that is, reduces to $J_{0,\sigma}$ as defined above for $\alpha=0$.

@@

@@

Then equation \ref{eqn:penderiv} can be re-written as
\[
  \langle D(T_{\sigma}[m]u)\delta m, T_{\sigma}[m]u\rangle = \langle DF[m]\delta m, 
\]
\begin{equation}
  \label{eqn:penderiv1}
   K[v_{0,\sigma}[m,u]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u]\rangle
\end{equation}
Combine equations \ref{eqn:resderiv} and \ref{eqn:penderiv1} with
equation \ref{eqn:basederiv} to obtain
\[
  \nabla \tilde{J}_{\alpha,\sigma}[m] =
  DF[m]^T\{ K[u_{\alpha,\sigma}[m]]^T (S[m]u_{\alpha,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:deriv}
  + \alpha^2 (
  K[v_{0,\sigma}[m, u_{\alpha,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+ K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{\alpha,\sigma}[m]]) \}.
\end{equation}

@@

The derivative of the expression on the right-hand side of
\ref{eqn:wg} is the sum of two terms, the first of which is given by
\ref{eqn:penderiv1} with $u = u_{0,\sigma}[m]$. The second is
\[
  \langle T_{\sigma}[m]u_{0,\sigma}[m], T_{\sigma}[m]D_m
  u_{0,\sigma}[m]\delta m \rangle
\]
which from identity \ref{eqn:blockderiv} is
\[
  = -\langle  T_{\sigma}[m]^TT_{\sigma}[m]u_{0,\sigma}[m], (S[m]^TS[m] + \sigma^2I)^{-1}
\]
\[
  \times ((DS[m]^{T}\delta m) (S[m]u_{0,\sigma}[m]-d) +
  S[m]^T(DS[m]\delta) m u_{0,\sigma}[m])
\]
Since
$T_{\sigma}[m]^TT_{\sigma}[m] = W(Ru_{0,\sigma}[m],-2,2)$, this is
\[
  = -\langle (DS[m]\delta m) (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m],
  S[m]u_{0,\sigma}[m] - d \rangle
\]
\begin{equation}
  \label{eqn:term2}
  - \langle S[m] (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m], (DS[m]\delta
  m)u_{0,\sigma}[m] \rangle
\end{equation}
Introduce the abbreviation
\begin{equation}
  \label{eqn:defv1}
  v_{1,\sigma}[m] = (S[m]^TS[m] + \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m].
\end{equation}
Then the RHS of equation \ref{eqn:term2bis} may be re-written as
\begin{equation}
  \label{eqn:term2bis}
  = -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle.
\end{equation}
Recall that we have just computed the second of two terms, the sum of
which is the derivative of $\tilde{J}_{\sigma}[m]$. The first is
\ref{eqn:penderiv1} with $u = u_{0,\sigma}[m]$. Combine the two
terms to obtain
\[
  D\tilde{J}_{\sigma}[m] \delta m = \langle DF[m]\delta m,
\]
\[
  K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{0,\sigma}[m]]\rangle
\]
\[
  -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle
\]
\[
  = \langle DF[m]\delta
  m,(K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]-K[v_{1,\sigma}[m]])^T(S[m]u_{0,\sigma}[m]-d)
\]
\[
  + K[u_{0,\sigma}[m]]^T S[m] (v_{0,\sigma}[m,u_{0,\sigma}[m]] -
  v_{1,\sigma}[m]) \rangle.
\]
To simplify this expression further, and reduce the computational
complexity of the derivative expression, note that the
operator-valued function $K$ is linear in its argument, according to
its definition \ref{eqn:convdef}. So the above is
\begin{equation}
  \label{eqn:wgderiv}
  = \langle DF[m]\delta m, (K[v_{\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d) 
  + K[u_{0,\sigma}[m]]^T S[m] v_{\sigma}[m] )\rangle.
\end{equation}
in which
\[
  v_{\sigma}[m] = v_{0,\sigma}[m,u_{0,\sigma}[m]] -  v_{1,\sigma}[m].
\]
The definitions \ref{eqn:defv0} and \ref{eqn:defv1} imply that
$v_{\sigma}[m]$ is the solution of the linear system
\begin{equation}
  \label{eqn:defvsigma}
  (S[m]^TS[m]+ \sigma^2I) v_{\sigma}[m] =
  (W(R(W(Ru_{0,\sigma}[m],-2,1)u_{0,\sigma}),2,0) - W(Ru_{0,\sigma}[m],-2,2))
  u_{0,\sigma}[m].
\end{equation}
That is, computation of $v_{\sigma}[m]$, hence of $D
\tilde{J}_{\sigma}[m]\delta m$, requires the solution of only one additional
linear system, beyond that necessary to compute the value of $\tilde{J}_{\sigma}[m]$.

@@

It remains to establish that the same result (expression
\ref{eqn:wgderiv}) for the derivative can be derived from the
limit \ref{eqn:wglim}. From from equation \ref{eqn:basederiv},
\[
  \frac{1}{\alpha^2}(D \tilde{J}_{\alpha,\sigma}[m] \delta m - D
  \tilde{J}_{0,\sigma}[m]\delta m) =
\]
\[
  = \frac{1}{\alpha^2} \left(\langle D_m(S[m]u)\delta m_{u =  u_{\alpha,\sigma}[m]}, (S[m]u_{\alpha,\sigma}[m]-d) \rangle\right.
\]
\[
  - \left.\langle D_m(S[m]u)\delta m_{u =  u_{0,\sigma}[m]}, (S[m]u_{0,\sigma}[m]-d) \rangle\right)
\]
\begin{equation}
  \label{eqn:limdiff}
  +  \langle D_m(T_{\sigma}[m]u)\delta
  m|_{u=u_{\alpha,\sigma}[m]},T_{\sigma}[m]u_{\alpha,\sigma}[m]\rangle_U
\end{equation}
Since the normal operator \ref{eqn:normal} is uniformly positive
definite hence boundedly invertible uniformly in $\alpha \ge 0$,
$u_{\alpha,\sigma}[m] \rightarrow u_{0,\sigma}[m]$ in $U$. All of the
operators appearing in the third term are $U-$continuous, so  the
third term tends to
\[
  \langle D_m(T_{\sigma}[m]u)\delta
  m|_{u=u_{0,\sigma}[m]},T_{\sigma}[m]u_{0,\sigma}[m]\rangle_U
\]
which is identixal to the left-hand side of equation
\ref{eqn:penderiv1} with $u=u_{0,\sigma}[m]$ and is the first term in the direct computation of
the $\tilde{J}_{\sigma}$ derivative, analyzed above.

It remains to understand
\[
\frac{1}{\alpha^2} \left(\langle D_m(S[m]u)\delta m_{u =  u_{\alpha,\sigma}[m]}, (S[m]u_{\alpha,\sigma}[m]-d) \rangle\right.
\]
\[
  - \left.\langle D_m(S[m]u)\delta m_{u =  u_{0,\sigma}[m]}, (S[m]u_{0,\sigma}[m]-d) \rangle\right)
\]
\[
  =\frac{1}{\alpha^2} \left(\langle D_m(S[m]u)\delta m_{u =  u_{\alpha,\sigma}[m]-u_{0,\sigma}[m]}, (S[m]u_{\alpha,\sigma}[m]-d) \rangle\right.
\]
\begin{equation}
  \label{eqn:resterms}
  + \left.\langle D_m(S[m]u)\delta m_{u =  u_{0,\sigma}[m]},
    (S[m](u_{\alpha,\sigma}[m]-u_{0,\sigma}[m]) \rangle\right)
\end{equation}
Observe that the normal equation \ref{eqn:normal} implies that
\[
  (S[m]^TS[m] + \sigma^2 I)(u_{\alpha,\sigma}[m]-u_{0,\sigma}[m]) =
  -\alpha^2 T_{\sigma}[m]^TT_{\sigma}[m]u_{\alpha,\sigma}[m]
\]
Since $u_{\alpha,\sigma}[m] \rightarrow u_{0,\sigma}$ in the sense of
$U$ as $\alpha \rightarrow 0$, it follows from definition \ref{eqn:defv1} that
\[
  \frac{1}{\alpha^2 }u_{\alpha,\sigma}[m]-u_{0,\sigma}[m] \rightarrow
  - v_{1,\sigma}[m], \, \alpha \rightarrow 0
\]
Writing as before $D_m(S[m]u)\delta m_{u =  ...} =
(DS[m]\delta m)...$,  the quantity on the right-hand side of equation
\ref{eqn:resterms}, obtain
\[
  \lim_{\alpha \rightarrow 0} 
\frac{1}{\alpha^2} \left(\langle D_m(S[m]u)\delta m_{u =  u_{\alpha,\sigma}[m]}, (S[m]u_{\alpha,\sigma}[m]-d) \rangle\right.
\]
\[
  - \left.\langle D_m(S[m]u)\delta m_{u =  u_{0,\sigma}[m]}, (S[m]u_{0,\sigma}[m]-d) \rangle\right)
\]
\[
  = - \langle (DS[m]\delta m)v_{1,\sigma}[m], S[m]u_{0,\sigma}[m]-d
  \rangle
\]
\[
  - \langle (DS[m]\delta m) u_{0,\sigma}[m],S[m]v_{1,\sigma}[m] \rangle
\]
\[
  =-\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle.
\]
which is precisely the right-hand side of equation
\ref{eqn:term2bis}. That is, the limit of the first two terms on the
right-hand side of equation \ref{eqn:limdiff} as $\alpha \rightarrow
0$ is exactly the second term of the direct derivative
computation. Since we have already shown that the third term in
\ref{eqn:limdiff} limits to
the first term in the direct computation, a second derivation of the
derivative $D\tilde{J}_{\sigma}[m]$ has been accomplished.

We have also shown that the limit of the scaled difference of the
derivatives is the derivative of the limit of the scaled difference,
pointwise in $m \in M$. In fact the limit is locally uniform in a
suitable sense, but that is a topic for another discussion.

\append{Unpreconditioned AWI: Matched Source Waveform Inversion}
Matched Source Waveform Inversion (MSWI) is a close relative to
AWI, differing in the use of a simpler penalty operator $T$, defined
by
\begin{equation}
  \label{eqn:upf}
  Tu (\bx_r,t;\bx_s) = t u(\bx_r,t;\bx_s).
\end{equation}
Similarly to AWI, $m$ is estimated by minimizing
\begin{equation}
  \label{eqn:junp}
 J^{\rm unp}_{\alpha,\sigma}[u,m,d] = \frac{1}{2}(\|S[m]u - d\|_D^2 + \alpha^2\|Tu\|_U^2 + \sigma^2 \|u\|^2_U).
\end{equation}
or equivalently
\begin{equation}
  \label{eqn:jtildeunp}
 \tilde{J}^{\rm unp}_{\alpha,\sigma}[m,d] = \frac{1}{2}(\|S[m]u_{\alpha,\sigma}[m,d] - d\|_D^2 + \alpha^2\|Tu_{\alpha,\sigma}[m,d]\|_U^2 + \sigma^2 \|u_{\alpha,\sigma}\|^2_U).
\end{equation}
in which $u_{\alpha,\sigma}$ is the solution of the normal equation
\begin{equation}
  \label{eqn:normalunp}
  (S[m]^TS[m] + \alpha^2T^TT + \sigma^2I)u_{\alpha,\sigma}[m] = S[m]^Td
\end{equation}
The only difference between AWI and MSWI lies in the defintiion of the
penalty operator.

MSWI was introduced in Hua Song's thesis
\cite[]{Song:94c} using a different definition of $T$ (but also
$m$-independent). The penalty operator defined in \ref{eqn:upf}, and
the associated inversion algorithm, was explored by Huang and Symes
\cite[]{HuangSymes2015SEG,HuangSymes:Geo17}, who show that at least to
some extent it avoids the cycle-skipping pathology of FWI. Note that
the division by $Ru_{0,\sigma}[m]$ in the definition of
$T_{\sigma}[m]$ is responsible for the global relation between AWI and
travel time tomography, within a limited class of problems, and this
relation cannot be asserted for MSWI.

Since the MSWI penalty operator $T$ is $m$-independent, it does not
appear explicilly in the the gradient
of the reduced MSWI objective $\tilde{J}^{\rm unp}_{\alpha,\sigma}$, as
given by the VPM gradient formula:
\[
  \nabla \tilde{J}^{\rm unp}_{\alpha,\sigma}[m] =
  D(S[m]u)^T_{u=u_{\alpha,\sigma}[m]}(S[m]u_{\alpha,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:tildejunpgrad}
  = DF[m]^T(K[u_{\alpha,\sigma}[m]]^T(S[m]u_{\alpha,\sigma}[m]-d).
\end{equation}

\append{Remarks on VCL Implementation}
\noindent {\bf 1.} $F$ is the mapping implemented (in finite difference approximation) in {\tt asg.fsbop}, with the additional constraint that the buoyancy field is fixed so that the computational version of ${\cal M}$ consists of (gridded) bulk modulus fields.

\noindent {\bf 2.} The convolution operator implemented in the module {\tt segyvc}
actually implements a time-discrete version of $K$ (or $L$). Note that
the transpose $K[u]^T$ is a restricted version of cross-correlation,
implemented as the transpose of {\tt segyvc.ConvolutionOperator}.

\noindent {\bf 3.} IWAVE implements the mappings $R$ and $W$ in the form of
the commands \\{\tt iwave/trace/main/rms.x} and {\tt
  iwave/trace/main/txgain.x}. The natural python implementation of $R$
would output the vector of trace norms as a python array (or a NumPy
ndarray). However that would require extracting the number of traces
from the input, and checking that the inputs are compatible. This is
already implicitly done in the IWAVE manipulations of SU trace
data. So {\tt iwave/trace/main/rms.x} stores the vector of trace norms
as the 0th data sample in a set of traces, one for each trace of the
input data: all header words are the same except for {\tt ns}, set to 1.

\noindent {\bf 4.} The python functions {\tt awi.setrms} and {\tt awi.applytxgain} call
the two IWAVE commands, manipulating the data as SU data files, rather
than as VCL vectors. That is, these are simply functions, not {\tt
  vcl.Function} objects. Since VCL vectors defined on a {\tt segyvc.Space}
have access to their data files (keyword {\tt data}), these are
convenient to use to express the action of $R$ and $W$ in AWI penalty operator,
gradient, and so on.

\noindent {\bf 5.} The implementation {\tt vcalg.conjgrad} of the Conjugate Gradient algorithm for the normal equations optionally returns the residual vector (parameter {\tt e}). The obvious way to use this is to create a 3-component product space for the output of the block-column operator $(S[m],\alpha T[m], \sigma I)^T$, and set the right-hand side vector (parameter {\tt b})
to $(d,0,0)^T$. Then on return one-half the norm-squared of {\tt e} is precisely the (approximate) value of $\tilde{J}_{\alpha,\sigma}[m]$.
